An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-28 paper/paper.tex
Paper: fix Intro based on comments from Mihaly

@@

Although they have not gained much popularity, peer to peer (P2P) file backup systems have been created and studied by researchers for over a decade. These systems do not suffer from the aforementioned drawbacks of centralized systems; they are free to use, limited in storage only by the user, and do not require a trusted central entity. Instead of backing up their data to a server, peers back up their data to \textit{each other}. However, abandoning the centralized server model in favor of a distributed one brings with it new problems that must be addressed before wide-spread adoption can be expected. Research in this area has focused on topics such as handling node churn \cite{ChurnResilient}, efficient routing algorithms \cite{Kademlia}, and minimizing bandwidth \cite{PeerStore}.

We have designed and implemented a new P2P backup system, called \mbox{BTBackup}, that improves the speed of backing up and recovering data. Our main contribution, and the technology responsible for faster file backup and recovery, is the way in which we handle data transfer: BitTorrent Sync, a new synchronization application created by BitTorrent, Inc. BitTorrent Sync uses a modified version of the BitTorrent protocol \cite{BTSyncFAQ} in order to quickly synchronize files between groups of peers. This protocol is especially effective at transferring large files, which makes its use perfect for the domain of P2P file backup. By leveraging BitTorrent Sync, we take advantage of BitTorrent's ability to better utilize a peer's bandwidth by downloading from multiple sources at once. This is in contrast to other P2P backup systems, which directly download from one source at a time. Our approach takes advantage of the fact that P2P backup systems require data replication, the act of inserting several redundant copies of data into the network, to combat node churn. We view the set of nodes storing replications for a given file as a BitTorrent swarm, allowing us to use the BitTorrent protocol without additional overhead.

We created test scenarios to (1) measure the extent to which it increases data backup and recovery transfer speeds, and (2) ensure our system is a feasible backup solution under realistic conditions. To test the former, we measured the rates of data transfer over time when backing up 1 GB of data. In order to see the effects filesize has on speed, we ran this test thrice: once with one 1GB file, once with ten 100MB files, and once with one hundred 10MB files. To test the latter, we created a test scenario that simulated usage of our system in a realistic environment. Each instance of this scenario created 100 peers on a network using MTA SZTAKI's cloud computing environment: SZTAKI Cloud. Specifically, this scenario measured the effects of node churn, or the rate at which nodes enter and leave the network, on our system. Since our system handles churn by moving replicas from unavailable nodes to available ones, we measure how much extra network traffic is generated to see the effects of churn. We take this measurement by running a test instance with no churn, and then running another one with churn. To provide a realistic setting, we had each peer limit their download and upload speeds to values that are based off of worldwide Internet bandwidth statistics.

After running our tests, we found that BitTorrent performs well as the data layer transport protocol for a P2P backup system. Specifically, we found that larger but fewer files are handled more efficiently than a greater number of smaller files, given the same amount of total data. We also determined that our system handles node churn well, when the rate of churn is modeled off of the average Internet user. Our final and most important result is that our system was able to achieve data transfer rates up to \~300\% quicker than the rates that could be realistically achieved in traditional backup systems.

@@

\subsection{Peer to Peer Networks}

@@

In this scenario, we tested how our system handles a constant amount of data spread out across a varying number of files. We run this test for three reasons. First, it allows us to test the capabilities of BitTorrent Sync in handling a large number of backups. From the data gathered here, it is possible to determine how well BitTorrent Sync would scale for use in a production system, where thousands of users could each be backing up thousands of files of varying sizes. Second, it allows us to measure the performance of BitTorrent against different file sizes. Since the use case for BitTorrent is typically downloading large files, it is important to determine whether or not our system will perform well when users try to backup many small files, should it be used in a production environment. Third, it will demonstrate the ability of our system to backup and recover files faster than if they were transferred by a direct download protocol.

@@

@@

@@

To test whether or not our system provided an increase in file transfer speeds, in addition to providing solutions for the other challenges of P2P backup systems, we implemented a simplified version of our system and tested it on MTA SZTAKI's cloud computing system. We found that it was able to handle node churn with a reasonable amount of overhead in network traffic. More importantly, however, we found that our system allowed for a \~300\% improvement in file backup and restoration speeds. This would not be possible if these data transfers were only between a user and some other entity, whether it is the service provider in a centralized backup system or a single peer in another P2P backup system.

>>>>

\hyphenation{MTA SZTAKI}

@@

Although they have not gained much popularity, peer to peer (P2P) file backup systems have been created and studied by researchers for over a decade. These systems do not suffer from the aforementioned drawbacks of centralized systems; they are free to use, limited in storage only by the user, and do not require a trusted central entity. Instead of backing up their data to a server, peers back up their data to \textit{each other}. However, abandoning the centralized server model in favor of a distributed one brings with it new problems that must be addressed before wide-spread adoption can be expected. Research in this area has focused on topics such as handling node churn\footnote{Node churn refers to the continuous arrival and departure of nodes in a P2P network. See Sections  \ref{subsec:PeertoPeerNetworks} and \ref{sec:churn} for more detail.} \cite{ChurnResilient}, efficient routing algorithms \cite{Kademlia}, and minimizing bandwidth \cite{PeerStore}.

We have designed and implemented a new P2P backup system, called \mbox{BTBackup}, that improves the speed of backing up and recovering data. Our primary contribution, and the technology responsible for faster file backup and recovery, is the way in which we handle data transfer: BitTorrent Sync, a new synchronization application created by BitTorrent, Inc. BitTorrent Sync uses a modified version of the BitTorrent protocol \cite{BTSyncFAQ} in order to quickly synchronize files between groups of peers. This protocol is especially effective at transferring large files, which makes its use perfect for the domain of P2P file backup. By leveraging BitTorrent Sync, we take advantage of BitTorrent's ability to better utilize a peer's bandwidth by downloading from multiple sources at once. This is in contrast to other P2P backup systems, which directly download from one source at a time. Our approach takes advantage of the fact that P2P backup systems require data replication, the act of inserting several redundant copies of data into the network, to combat node churn. We view the set of nodes storing replications for a given file as a BitTorrent swarm, allowing us to use the BitTorrent protocol without additional overhead.

A secondary contribution of our research is the testing infrastructure we have designed. We created test scenarios to (1) measure the extent to which BitTorrent increases data backup and recovery transfer speeds, and (2) ensure our system is a feasible backup solution under realistic conditions. To test the former, we measured the rates of data transfer over time when backing up 1 GB of data. In order to see the effects filesize has on speed, we ran this test thrice: once with one 1GB file, once with ten 100MB files, and once with one hundred 10MB files. To test the latter, we created a test scenario that simulated usage of our system in a realistic environment. Each instance of this scenario created 100 peers on a network using MTA SZTAKI's cloud computing environment: SZTAKI Cloud. Specifically, this scenario measured the effects of node churn, or the rate at which nodes enter and leave the network, on our system. Since our system handles churn by moving replicas from unavailable nodes to available ones, we measure how much extra network traffic is generated to see the effects of churn. We take this measurement by running a test instance with no churn, and then running another one with churn. To provide a realistic setting, we had each peer limit their download and upload speeds to values that are based off of worldwide Internet bandwidth statistics.

After running our tests, we found that BitTorrent performs well as the data layer transport protocol for a P2P backup system. Specifically, we found that \emph{larger but fewer files are handled more efficiently than a greater number of smaller files}, given the same amount of total data. We also determined that our system handles node churn well, when the rate of churn is modeled off of the average Internet user. Our final and most important result is that \emph{our system was able to achieve data transfer rates up to \~300\% quicker than the rates that could be realistically achieved in traditional backup systems.}
% TODO emphasize that our system handles node churn well, but need to reword/give numeric data
% TODO add (first, create) result of file backup speed increase percentage

@@

\subsection{Peer to Peer Networks} \label{subsec:PeertoPeerNetworks}

@@

% TODO clarify that we are actually only testing 1 variable here, chunk size
In this scenario, we tested how our system handles a constant amount of data spread out across a varying number of files. We run this test for three reasons. First, it allows us to test the capabilities of BitTorrent Sync in handling a large number of backups. From the data gathered here, it is possible to determine how well BitTorrent Sync would scale for use in a production system, where thousands of users could each be backing up thousands of files of varying sizes. Second, it allows us to measure the performance of BitTorrent against different file sizes. Since the use case for BitTorrent is typically downloading large files, it is important to determine whether or not our system will perform well when users try to backup many small files, should it be used in a production environment. Third, it will demonstrate the ability of our system to backup and recover files faster than if they were transferred by a direct download protocol.

@@

In addition to giving us insight into the effects file size and the number of files has on the speed of data backup, we used this test scenario to determine the extent that BitTorrent increases the data transfer rate of file backup and recovery.

@@

% TODO rename this test to 'Data Backup Speed From Peer's Perspective' or something, discuss increase in backup speeds (upload speed of peer vs upload speed of peer + upload speed of nodes. need to do algebra to figure out speed increase)

@@

To test whether or not our system provided an increase in file transfer speeds, in addition to providing solutions for the other challenges of P2P backup systems, we implemented a prototype of our system and tested it on a virtualized, 100 node network using MTA SZTAKI's cloud computing system. We found that it was able to handle node churn with a reasonable amount of overhead in network traffic, requiring the replacement of a replica every 11 days. More importantly, however, we found that our system allowed for a \~300\% improvement in file restoration speeds. This would not be possible if these data transfers were only between a user and some other entity, whether it is the service provider in a centralized backup system or a single peer in another P2P backup system.
% TODO add (first, create) result of file backup increase
